---
title: "Analysis of Factors Affecting Shots in the NBA"
output: github_document
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 4, echo = FALSE)
```

## Group :

Llewellyn Curtis\
Henry Earl\
Vaughan Boniface\
Henry Edgington\
Fintan Pawley\
Scott Airey

\newpage

## Abstract:

This report examines the factors influencing shot success in the NBA, with a primary focus on the spatial relationship between shooters and defenders. Utilizing player tracking data, we analyze the frequency distribution of defender proximity and its impact on shot conversion. We then also look broadly at the nature of shooting in the NBA and whether shot accuracy or frequency play a bigger role in the game outcome.

## Section 1: Proximity of Defenders

This analysis explores the dynamics of shot success, specifically focusing on the distance relationship between shooters and defenders. While intuition suggests that a tighter gap between players would lead to lower shot percentages, we can look a bit deeper at nature of the defensive player to observe trends.

### Defender Distance

The section begins by mapping the frequency distribution of defender distances to understand the typical shooting senarios we can then look into. Figure 1 visualizes the frequency distribution of the closest defender, illustrating how many shot attempts occur at specific distances, in metres, from the nearest defender.

```{r}
NBA <- read.csv("NBA_Shots.csv", header=TRUE, stringsAsFactors=FALSE)
def_dist_table <- table(NBA$CLOSE_DEF_DIST)
barplot(def_dist_table,main = "Figure 1: Frequency Distribution of Defender Distance", xlab = 'Closest Defender Distance', ylab = 'Frequency')
```

The plot shows that the data is heavily left-skewed, with a distinct peak occurring between roughly 2 and 6 feet. This indicates that the vast majority of NBA shots are contested within a relatively tight window, reflecting the high level of defensive pressure. Beyond 10 feet, the frequency drops off significantly, representing "wide open" scenarios or transition plays where the defender has been beaten or lost in rotation. Notably, the presence of data points extending as far as 40+ feet highlights the "messy" nature of raw NBA data. This distribution serves as the baseline for our analysis, allowing us to identify the primary clusters of play before filtering out these extreme outliers to find a more accurate correlation between proximity and shot success.

### Shot Conversion

This bar plot visualizes the relationship between the distance of the closest defender and the probability of a shot being successful. It represents a refined iteration of the previous dataset, specifically filtered to exclude low-frequency outliers, to provide a less volatile trend of how defensive proximity impacts shot conversion.

```{r}
NBA_3pt <- subset(NBA, PTS_TYPE == 3)
test_table_3pt <- table(NBA_3pt$CLOSE_DEF_DIST, NBA_3pt$FGM)
test_table_3pt_clean <- test_table_3pt[apply(test_table_3pt != 0, 1, all),]
success_probs_3pt_clean <- prop.table(test_table_3pt_clean, 1)[,2]
barplot_3pt_clean <- barplot(success_probs_3pt_clean, main = "Figure 2: Section of Defender Distance", ylim=c(0,1), ylab='Probability of Shot Going in', xlab='Closest Defender Distance', xaxs = 'i')
```

The bar plot reveals a clear upward trend in scoring probability as the distance of the closest defender increases, particularly within the 0 to 10-foot range. While the likelihood of a successful shot generally improves with more space, the data becomes significantly more volatile beyond 15 feet. Even after removing low-frequency occurrences to clean the "messy" numbers, the erratic spikes at greater distances suggest that these wide-open scenarios are less common, causing individual makes or misses to have a disproportionate impact on the calculated probability. Ultimately, the graph demonstrates that while increased distance generally aids the shooter, the relationship is most consistent when defenders are in closer proximity.

### Player Heights

Building on the previous findings, this section examines whether the physical stature of the nearest Defender influences shooting conversion. Defenders will be categorized into short, medium, and tall groups based on height quartiles. By comparing these specific subgroups against the general trends identified earlier, we can determine if a defender’s height acts as a multiplier that offsets the advantage of offensive space.

```{r}
NBA_players <- read.csv("Player_Info.csv", header=TRUE, stringsAsFactors=FALSE)
NBA_players$PLAYER_NAME <- paste(NBA_players$First_Name, NBA_players$Surname)
short <- subset(NBA_players, Height <= 190)
medium <- subset(NBA_players, Height < 205 & Height > 190)
tall <- subset(NBA_players, Height >= 205)
short_shots <- subset(NBA, CLOSEST_DEFENDER %in% short$PLAYER_NAME)
medium_shots <- subset(NBA, CLOSEST_DEFENDER %in% medium$PLAYER_NAME)
tall_shots <- subset(NBA, CLOSEST_DEFENDER %in% tall$PLAYER_NAME)

short_shots_3pt <- subset(short_shots, PTS_TYPE == 3)
short_shots_3pt_table <- table(short_shots_3pt$CLOSE_DEF_DIST, short_shots_3pt$FGM)
table_short_3pt_clean <- short_shots_3pt_table[apply(short_shots_3pt_table != 0 & short_shots_3pt_table != 1, 1, all),]
success_probs_short_3pt_clean <- prop.table(table_short_3pt_clean, 1)[,2]

par(mfrow = c(1, 3), oma = c(0, 0, 3, 0))
barplot_short_3pt <- barplot(success_probs_short_3pt_clean, ylim=c(0,1), ylab='Probability of Shot Going in', xlab='Closest Defender Distance', main = 'Short Defenders', xaxs = 'i')

medium_shots_3pt <- subset(medium_shots, PTS_TYPE == 3)
medium_shots_3pt_table <- table(medium_shots_3pt$CLOSE_DEF_DIST, medium_shots_3pt$FGM)
table_medium_3pt_clean <- medium_shots_3pt_table[apply(medium_shots_3pt_table != 0 & medium_shots_3pt_table != 1, 1, all),]
success_probs_medium_3pt_clean <- prop.table(table_medium_3pt_clean, 1)[,2]
barplot_medium_3pt <- barplot(success_probs_medium_3pt_clean, ylim=c(0,1), ylab='Probability of Shot Going in', xlab='Closest Defender Distance', main = 'Medium Defenders', xaxs = 'i')

tall_shots_3pt <- subset(tall_shots, PTS_TYPE == 3)
tall_shots_3pt_table <- table(tall_shots_3pt$CLOSE_DEF_DIST, tall_shots_3pt$FGM)
table_tall_3pt_clean <- tall_shots_3pt_table[apply(tall_shots_3pt_table != 0 & tall_shots_3pt_table != 1, 1, all),]
success_probs_tall_3pt_clean <- prop.table(table_tall_3pt_clean, 1)[,2]
barplot_tall_3pt <- barplot(success_probs_tall_3pt_clean, ylim=c(0,1), ylab='Probability of Shot Going in', xlab='Closest Defender Distance', main = 'Tall Defenders', xaxs = 'i')

mtext("Figure 3: Impact of Defender Height on 3pt Shot Success", 
      outer = TRUE, 
      cex = 1.5, 
      font = 2)
```

Upon reviewing the individual distributions for "Short," "Medium," and "Tall" players in figure 3, it is evident that all three groups exhibit a broadly similar trend: shot probability scales positively with defender distance. Regardless of the shooter's physical stature, the "decay" in shooting efficiency under tight defensive pressure remains the same.

```{r}
short_bs <- aggregate(FGM ~ CLOSE_DEF_DIST, data = short_shots_3pt, FUN = mean)
short_probs_bs <- short_bs[apply(short_bs != 0 & short_bs != 1, 1, all),]

medium_bs <- aggregate(FGM ~ CLOSE_DEF_DIST,data = medium_shots_3pt, FUN = mean)
medium_probs_bs <- medium_bs[apply(medium_bs != 0 & medium_bs != 1, 1, all),]

tall_bs <- aggregate(FGM ~ CLOSE_DEF_DIST,data = tall_shots_3pt, FUN = mean)
tall_probs_bs <- tall_bs[apply(tall_bs != 0 & tall_bs != 1, 1, all),]

og_bs <- aggregate(FGM ~ CLOSE_DEF_DIST, data = NBA_3pt, FUN = mean)
og_probs_bs <- og_bs[apply(og_bs != 0 & og_bs != 1, 1, all),]

plot(short_probs_bs, main = "Figure 4: Comparing Player Heights", type = 'n',ylab='Probability of Shot Going in', xlab='Closest Defender Distance')

lines(short_probs_bs, col = c('gold'), lwd = 2)
lines(medium_probs_bs, col = c('blue'), lwd = 2)
lines(tall_probs_bs, col = c('forest green'), lwd = 2)
lines(og_probs_bs, col = c('red'), lwd = 2)

model_short <- lm(short_probs_bs[,2] ~ short_probs_bs[,1])
abline(model_short, col = "gold", lwd = 4)

model_mid <- lm(medium_probs_bs[,2] ~ medium_probs_bs[,1])
abline(model_mid, col = "blue", lwd = 4)

model_tall <- lm(tall_probs_bs[,2] ~ tall_probs_bs[,1])
abline(model_tall, col = "forest green", lwd = 4)

model_og <- lm(og_probs_bs[,2] ~ og_probs_bs[,1])
abline(model_og, col = "red", lwd = 4)


legend('topright', legend = c('Short', 'Medium', 'Tall', 'Average'), col = c('gold', 'blue', 'forest green', 'red'), lty = c(1,1,1,1), lwd = 2)
```

The visual comparison confirms that player height does not fundamentally alter the relationship between defender proximity and shot success. While one might hypothesize that taller players that can get over attackers could mitigate the impact of close-range defenders, the data suggests that the physiological advantage of height is secondary to the physical disruption caused by defensive "crowding."

In all three categories, the highest concentration of high-probability shots occurs when the nearest defender is beyond 4 meters. Consequently, for the remainder of this analysis, we can treat "defender distance" as a universal variable, regardless of the individual physical profiles of the players involved.

\newpage

### Closest Defender Role

Now, we will look at how a player's position affects their shot defending capabilities. This section explores whether a defender's primary role—either as a "True Defender" or an "Attacker Defender" changes the probability of an opponent making a shot. By examining the data in figure 5, we can identify specific patterns in how different types of players contest shots at various distances.

```{r}
par(mfrow = c(1, 3), oma = c(0, 0, 3, 0))
NBA_combined <- merge(NBA, NBA_players, by = c('PLAYER_NAME'))
defenders <- subset(NBA_combined, Pos == 'PG' | Pos == 'SG')
true_defenders <- subset(NBA_combined, CLOSEST_DEFENDER %in% defenders$PLAYER_NAME)
TD_table <- table(true_defenders$CLOSE_DEF_DIST, true_defenders$FGM)
TD_table_clean <- TD_table[apply(TD_table != 0 & TD_table != 1, 1, all),]
TD_probs <- prop.table(TD_table_clean, 1)[,2]
barplot(TD_probs,ylab='Probability of Shot Going in', xlab='Closest Defender Distance', main = 'True Defenders', xaxs = 'i')
attackers <- subset(NBA_combined, Pos == 'PF' | Pos == 'C')
attack_defenders <- subset(NBA_combined, CLOSEST_DEFENDER %in% attackers$PLAYER_NAME)
AD_table <- table(attack_defenders$CLOSE_DEF_DIST, attack_defenders$FGM)
AD_table_clean <- AD_table[apply(AD_table != 0 & AD_table != 1, 1, all),]
AD_probs <- prop.table(AD_table_clean, 1)[,2]
barplot(AD_probs,ylab='Probability of Shot Going in', xlab='Closest Defender Distance', main = 'Attacker Defenders', xaxs = 'i')
defender_bs <- aggregate(FGM ~ CLOSE_DEF_DIST,data = true_defenders, FUN = mean)
def_prob_bs <- defender_bs[apply(defender_bs != 0 & defender_bs != 1, 1, all),]


attacker_bs <- aggregate(FGM ~ CLOSE_DEF_DIST,data = attack_defenders, FUN = mean)
attack_prob_bs <- attacker_bs[apply(attacker_bs != 0 & attacker_bs != 1, 1, all),]

plot(def_prob_bs, type = 'n', main='Comparison', ylab = 'Probability of Shot Going in', xlab='Closest Defender Distance')

lines(def_prob_bs, col = c('dark green'), lwd = 2)
lines(attack_prob_bs, col = c('maroon'), lwd = 2)

modeldef <- lm(def_prob_bs[,2] ~ def_prob_bs[,1])
abline(modeldef, col = "dark green", lwd = 4)
modelatt <- lm(attack_prob_bs[,2] ~ attack_prob_bs[,1])
abline(modelatt, col = "maroon", lwd = 4)

legend('bottomright', legend = c('Attacker Defenders', 'True Defenders'), col = c('dark green', 'maroon'), lwd = 3)

mtext("Figure 5: Shot Conversion Relative to Defensive Player Role", 
      outer = TRUE, 
      cex = 1.5, 
      font = 2)
```

The trends visualized show a clear distinction in defensive skills between the two groups. While "True Defenders" stay relatively consistent regardless of distance, the data indicates that attacking players are actually less effective when guarding at a close distance compared to defenders. However, as the distance increases, the attacking players' defensive impact improves, eventually performing better at a distance than the guard group. This suggests that while defenders have a more balanced skill set, the attacking players have a specific defensive range where their effectiveness peaks.

## Section 2: Player Positions

For the next analysis, we investigate the role that a player’s position may have on their accuracy when taking a shot. Since each position has their own rough area of the court they generally operate in, we can hypothesise that as a player’s shooting distance varies, there should be a correlation between the position they play and their accuracy at different ranges. Our hypotheses here are as follows:

**Point Guards** often play from beyond the three-point line while attacking and are generally the furthest players from the rim, as such they should be more comfortable with longer shots than other positions as they are more often in those long shot scenarios.

**Shooting Guards** are the second furthest out as they generally play along ‘the wings’ of the court at either the far left or far right sides so are often at three-point range, as such we are expecting them to have the second-best accuracy at longer ranges behind PGs.

**Power Forwards** are next and as their role tends to be more mid to close ranged within the two-point zone, they are expected to be noticeably stronger with close range shots than the PGs and SGs but may struggle a little more as the distance from the rim increases.

The final position we will look at are the **Centers**, they operate almost exclusively at close range, deep within the two-point area while attacking and as such we are surmising that they will be significantly stronger than the other positions when near to the rim. This also means that we are hypothesising that they are going to be significantly weaker at range than their teammates.

We aren’t analysing **Small Forwards** here as their position isn’t as strictly conforming to specific ranges as the others, so their data has been considered irrelevant for this analysis. To summarise, we are hypothesising that PGs and SGs will be more accurate with three pointers while Cs and PFs will be more accurate with two pointers as we look closer to the rim. It should also be noted that we are assuming traditional roles for the positions here and that there are outliers in the NBA due to the fluidity of its positions.

### Shooting Frequency of Different Player Roles

```{r}
NBA_players$PLAYER_NAME <- paste(NBA_players$First_Name, NBA_players$Surname)
NBA_combined <- merge(NBA, NBA_players, by = c('PLAYER_NAME'))

combined_table <- table(NBA_combined$PTS_TYPE, NBA_combined$Pos)
barplot(combined_table, main = "Figure 6: Shot Frequency per Position", legend = c('2 Points', '3 Points'), col = c('pink', 'brown'), beside = TRUE, ylab='Shot Frequency',xlab='Position')
```

In figure 6 we are able to see the frequencies of each positions shot types. So far our hypothesis for Centers remains accurate as they have taken far more two point than three-point shots, the same general trend is present for Power Forwards and the Point Guards and Shooting Guards have a higher frequency of three-point attempts than the PFs and Cs.

### Shoot Accuracy of Different Player Roles

```{r}
PG <- subset(NBA_combined, Pos == 'PG')

PG_table <- aggregate(FGM ~ SHOT_DIST, data = PG, FUN = mean)
PG_probs <- PG_table[apply(PG_table != 0 & PG_table != 1, 1, all),]

SG <- subset(NBA_combined, Pos == 'SG')

SG_table <- aggregate(FGM ~ SHOT_DIST, data = SG, FUN = mean)
SG_probs <- SG_table[apply(SG_table != 0 & SG_table != 1, 1, all),]

PF <- subset(NBA_combined, Pos == 'PF')

PF_table <- aggregate(FGM ~ SHOT_DIST, data = PF, FUN = mean)
PF_probs <- PF_table[apply(PF_table != 0 & PF_table != 1, 1, all),]

C <- subset(NBA_combined, Pos == 'C')

C_table <- aggregate(FGM ~ SHOT_DIST, data = C, FUN = mean)
C_probs <- C_table[apply(C_table != 0 & C_table != 1, 1, all),]

plot(PG_probs, main = "Figure 7: Shot Conversion at Distance", type = 'n',ylab='Probability of Shot Going in', xlab='Shot Distance'  )

lines(PG_probs, col = c('gold'), lwd = 2)
lines(SG_probs, col = c('red'), lwd = 2)
lines(PF_probs, col = c('forest green'), lwd = 2)
lines(C_probs, col = c('blue'), lwd = 2)

modelPG <- lm(PG_probs[,2] ~ PG_probs[,1])
abline(modelPG, col = "gold", lwd = 4)

modelSG <- lm(SG_probs[,2] ~ SG_probs[,1])
abline(modelSG, col = "red", lwd = 4)

modelPF <- lm(PF_probs[,2] ~ PF_probs[,1])
abline(modelPF, col = "forest green", lwd = 4)

modelC <- lm(C_probs[,2] ~ C_probs[,1])
abline(modelC, col = "blue", lwd = 4)

legend('topright', legend = c('PG', 'SG', 'PF', 'C'), col = c('gold', 'red', 'forest green', 'blue'), lty = c(1,1,1,1), lwd = 2)
```

When plotting our data into figure 7, we are able to observe that our original hypotheses were partially accurate. The Centers average accuracy is significantly better at within \~5 feet of the rim than the other positions and drops below them at \~20 feet. This could be a consequence of a lack of variation in their chosen position, perhaps in staying within their positions traditional area they’ve weakened their capability beyond it. The same can be seen in the results for the Shooting Guards and Point Guards as their data suggests a poorer performance at ranges within \~15 feet when compared to Power Forwards and Centers. Interestingly however, PFs, PGs, and SGs all converge to a similar point towards the extent of the distance. One could presume this to be due to the overall standard of ability within the NBA, perhaps three pointers are harder to defend against. If true then since they’re all professional basketball players the added distance could end up being negligible towards their accuracy.

Generally, the data indicates that our original hypotheses are accurate and we are able to draw some conclusions from these findings to apply to some real-world settings. The Centers results demonstrate that they are the strongest shot takers within 5 feet of the rim. For attacking plays, this means teams should prioritize getting the ball to their Center when in a close range position. Defensively, the opposing Center should be the top priority to stop. The convergence of PGs, PFs, and SGs at longer ranges suggests that they’re more interchangeable at a three-point range, this could be used to create more three-point chances as the PFs don’t necessarily need to stick to their traditional area when further out. The Shooting Guards having very similar three-point range accuracy to PGs and PFs suggests a lack of advantage from playing on ‘the wings’ of the court, this puts into question the value of SGs as a unique position from PGs and PFs. Considering this it could be worth teams choosing to allocate less salary towards traditional SGs, instead putting the money towards more versatile SGs.

\newpage

## Section 3: Shot Accuracy Vs Frequency

In this section, we will explore the relationship between how often shots are taken and how often they actually result in points. By analyzing the volume of attempts alongside scoring percentages, we can determine whether taking more shots necessarily leads to better offensive efficiency or if a high frequency eventually leads to diminishing returns in accuracy.

### Shot Accuracy

To begin this analysis, we will first look at how many of the taken shots go in for winning teams versus losing teams. We can compare shot success rates to see if there is a significant difference in conversion efficiency based on the final game outcome.

```{r}
df <- NBA

# 2. Filter the data for wins and losses and find individual accuracy
winning_games <- subset(df, W == "W")
losing_games <- subset(df, W == "L")

results_win <- aggregate(FGM ~ GAME_ID, 
                         data = winning_games, 
                         FUN = function(x) c(sum = sum(x), count = length(x)))
results_lose <- aggregate(FGM ~ GAME_ID, 
                          data = losing_games, 
                          FUN = function(x) c(sum = sum(x), count = length(x)))


# Create dataset to plot from above functions
results_win <- do.call(data.frame, results_win)
names(results_win) <- c("GAME_ID", "successful_shots", "total_shots")

results_lose <- do.call(data.frame, results_lose)
names(results_lose) <- c("GAME_ID", "successful_shots", "total_shots")

results_win$accuracy_rate <- results_win$successful_shots / results_win$total_shots
results_lose$accuracy_rate <- results_lose$successful_shots / results_lose$total_shots

density_wins <- density(results_win$accuracy_rate)
density_losses <- density(results_lose$accuracy_rate)

# Plot limits for x and y
x_limit <- range(c(density_wins$x, density_losses$x))
y_limit <- range(c(density_wins$y, density_losses$y))

# Create the initial plot for 'Wins'
plot(density_wins,
     main = "Figure 8: Shot Conversion by Game (Win/Loss)",
     xlab = "Shot Conversion",
     ylab = "Density",
     xlim = x_limit,
     ylim = y_limit,
     col = "blue",
     lwd = 2)

# Overlay the density for 'Losses'
lines(density_losses, col = "red", lwd = 2)

# Fill the area under the curves
polygon(density_wins, col = rgb(0, 0, 1, 0.3), border = NA)
polygon(density_losses, col = rgb(1, 0, 0, 0.3), border = NA)

# Add vertical lines for the mean for Wins and Losses
mean_win <- mean(results_win$accuracy_rate)
mean_lose <- mean(results_lose$accuracy_rate)

abline(v = mean_win, col = "darkblue", lty = 2, lwd = 2)
abline(v = mean_lose, col = "darkred", lty = 2, lwd = 2)
# Add a legend
legend("topleft",
       legend = c("Wins", "Losses", "Mean Conversion Rate (Wins)", "Mean Conversion Rate (Losses)"),
       col = c("blue", "red", "darkblue", "darkred"),
       lty = c(1, 1, 2, 2),
       lwd = c(2, 2, 2, 2),
       fill = c(rgb(0, 0, 1, 0.3), rgb(1, 0, 0, 0.3), NA, NA),
       border = NA,
       bg = "white")

```

As shown in figure 8, we can see that the average shot conversion in a game for a winning team is consistently higher. The density plot illustrates a clear separation between the two groups, with the mean conversion rate for wins sitting significantly to the right of the mean for losses. This suggests that while both winning and losing teams can have overlapping conversion percentages, maintaining a higher efficiency is a strong predictor of success. Ultimately, the data confirms that shot quality and successful conversion are key differentiators in determining the final outcome of a game.

### Shot Frequency

Following the analysis of shot conversion, we turn our attention to the volume of attempts made during a game. This sub-section examines the distribution of shot frequency to determine if winning is simply a matter of taking more shots, or if the total number of attempts remains consistent regardless of the game's outcome. We can contrast the sheer number of opportunities created by winning and losing teams to see how volume correlates with the efficiency metrics previously discussed.

```{r}
# Same again but sum shots
density_wins_vol <- density(results_win$total_shots)
density_losses_vol <- density(results_lose$total_shots)

# 3. Determine plot limits
x_limit_vol <- range(c(density_wins_vol$x, density_losses_vol$x))
y_limit_vol <- range(c(density_wins_vol$y, density_losses_vol$y))

# 4. Create the plot
plot(density_wins_vol,
     main = "Figure 9: Total Shots by Game (Win/Loss)",
     xlab = "Total Shot Attempts",
     ylab = "Density",
     xlim = x_limit_vol,
     ylim = y_limit_vol,
     col = "blue",
     lwd = 2)

# Overlay Losses
lines(density_losses_vol, col = "red", lwd = 2)

# Fill areas
polygon(density_wins_vol, col = rgb(0, 0, 1, 0.3), border = NA)
polygon(density_losses_vol, col = rgb(1, 0, 0, 0.3), border = NA)

# 5. Add vertical lines for Average Shot Volume
mean_vol_wins <- mean(results_win$total_shots)
mean_vol_lose <- mean(results_lose$total_shots)

abline(v = mean_vol_wins, col = "darkblue", lty = 2, lwd = 2)
abline(v = mean_vol_lose, col = "darkred", lty = 2, lwd = 2)

# 6. Add the legend
legend("topleft", 
       legend = c("Wins", "Losses", "Avg Shots (Wins)", "Avg Shots (Losses)"),
       col = c("blue", "red", "darkblue", "darkred"),
       lty = c(1, 1, 2, 2),
       lwd = 2,
       fill = c(rgb(0, 0, 1, 0.3), rgb(1, 0, 0, 0.3), NA, NA),
       border = NA,
       bg = "white")
```

As we can see in figure 9, the total number of shots taken is broadly the same for both winning and losing teams. The density curves overlap almost entirely, and the average shot attempts for both groups sit very close to one another on the x-axis. This indicates that simply increasing the volume of shots does not provide a guaranteed advantage, as losing teams often generate a similar number of opportunities as the winners.

### Comparison

When comparing frequency and accuracy, it becomes clear that shot conversion is what really matters more for securing a win. While the number of attempts remains consistent across outcomes, the efficiency with which those shots are made is what actually separates the two groups. This suggests that offensive success is driven by the quality of shots taken rather than the sheer quantity, making conversion rate the more critical metric for predicting performance.

```{r}
# Accuracy Gap
acc_diff <- (mean(results_win$accuracy_rate) - mean(results_lose$accuracy_rate)) * 100

# Volume Gap
vol_diff <- mean(results_win$total_shots) - mean(results_lose$total_shots)

cat("Winners shoot", round(acc_diff, 2), "% better than losers.\n")
cat("Winners take", round(vol_diff, 2), "more shots than losers.\n")
```

The raw data presented provides concrete evidence for this conclusion. While winners only take 1.24 more shots than losers on average, they shoot 5.12% better from the field.

Given that the average number of shots per game is relatively high, an extra shot or two is only a small increase in volume. However, a 5.12% difference in shot conversion is a notable margin that consistently translates into more points, confirming that accuracy is broadly more important than frequency in determining a win.

```{r}
results_win$status <- "Win"
results_lose$status <- "Loss"
combined_results <- rbind(results_win, results_lose)

plot_colors <- ifelse(combined_results$status == "Win", 
                      rgb(0, 0, 1, 0.5),
                      rgb(1, 0, 0, 0.5))

# Create scatter plot
plot(combined_results$total_shots, combined_results$accuracy_rate,
     main = "Figure 10: Accuracy Vs Frequency of Shots (Win/Loss)",
     xlab = "Total Shots Taken",
     ylab = "Field Goal Percentage (Accuracy)",
     pch = 19, 
     col = plot_colors)

# Trend lines for both groups
abline(lm(accuracy_rate ~ total_shots, data = results_win), 
       col = "darkblue", lwd = 2)
abline(lm(accuracy_rate ~ total_shots, data = results_lose), 
       col = "darkred", lwd = 2)

legend("topright", 
       legend = c("Wins", "Losses", "Trend: Wins", "Trend: Losses"),
       col = c("blue", "red", "darkblue", "darkred"),
       pch = c(19, 19, NA, NA),
       lty = c(NA, NA, 1, 2),
       lwd = 2,
       bg = "white")
```

In summary, the relationship between accuracy and frequency shown in figure 10 reinforces that offensive success is not built on volume alone. Since winning teams convert at a higher rate despite taking a similar number of shots as losing teams, it is clear that shot quality is the deciding factor. The data suggests that a medium amount of high-quality shots is the optimal strategy for winning games, as efficiency remains the primary driver of the final score.

## Conclusion:

This report demonstrates that NBA shot success is primarily dictated by the spatial relationship between shooters and defenders, rather than the physical stature of the players involved. Most shot attempts occur within a contested window of 2 to 6 feet, where defensive pressure is highest. While scoring probability increases as defenders drop back, this trend remains consistent regardless of whether a defender is short, medium, or tall, suggesting that the psychological and physical disruption of "crowding" is the dominant defensive factor. Furthermore, positional analysis reveals that while Centers dominate efficiency within 5 feet of the rim, the accuracy of Point Guards, Shooting Guards, and Power Forwards converges at longer ranges. This indicates that traditional roles are becoming increasingly interchangeable beyond the three-point line, providing teams with more flexibility in shot creation.

The final analysis of game outcomes reveals that offensive quality is a far more significant predictor of victory than the sheer volume of shots taken. Winning and losing teams generate a nearly identical number of opportunities, with winners taking only 1.24 more shots on average per game. However, winning teams maintain a 5.12% higher field goal percentage, proving that the ability to convert high-quality looks is the primary differentiator in the league. Ultimately, offensive success is not built on frequency alone; rather, it is the result of a strategic emphasis on shot quality and conversion efficiency over the quantity of attempts.
